\chapter{Build Systems} \label{buildsystems}

\section{The Yocto Project}

The Yocto Project is an open source collaborative project that provides users with a set of tools to create custom Linux-based systems for embedded products. It's based on the OpenEmbedded framework and is backed by the Linux Foundation. The Yocto Project works with hardware vendors, open source communities, and hundreds of developers to provide a robust development environment for embedded products.

Yocto Project allows developers to create unique Linux-based systems for embedded devices. Yocto Project provides developers with the tools to customise their embedded Linux systems to meet the specific needs of their products. Yocto Project is used by many companies for their embedded products. It is especially useful for those developing custom embedded products, as it allows users to quickly create a customised Linux-based operating system.

Yocto Project provides many features that make it a great choice for embedded Linux development. These features include:

\begin{enumerate}
	\item \textbf{Open Source} Yocto Project is an open source project backed by the Linux Foundation. This means it is free to use and developers can access the source code to customise their systems as needed
	\item \textbf{Compatibility} Yocto Project is compatible with many types of embedded hardware, including ARM, PowerPC, MIPS, and x86. This makes it easy to use for any type of embedded project
	\item \textbf{Robust Development Environment} Yocto Project provides a robust development environment for embedded Linux development. It includes libraries, tools, and debugging support to make development easier
	\item \textbf{High Performance} Yocto Project provides an optimised development environment for embedded systems. This helps developers to create high-performance products quickly and easily
	\item \textbf{Flexibility} Yocto Project provides developers with the flexibility to create custom Linux-based systems for their embedded devices. This allows developers to tailor their systems to meet the specific needs of their products
	\item \textbf{Time Savings} Yocto Project makes it easier and faster to create custom Linux-based systems. This helps to reduce development time and save money
\end{enumerate}